Stability in human interaction networks: primitive typology of vertex, prominence of measures and time activity statistics

### 2015-04-08 paper.tex
atualizando para repassar

\subsection{Prevalence of centrality over symmetry and symmetry over clusterization}\label{prevalence}
We have analyzed the topology of the email networks using standard, well-established metrics based on centrality and clustering.
We also introduced metrics related to symmetry because there is evidence that such metrics are relevant in a social context~\cite{newmanEvolving}.
The contribution from the distinct metrics to network topology is very similar for all the networks considered, and did not vary with time.
This stability in network behavior is remarkable, as can be noticed by the small standard deviations of the contribution of each metric to principal components (from usual Principal Component Analysis~\cite{pca}) along time of Tables~\ref{compPCA0}-\ref{compPCA2}.

@@

Degree and strength are highly correlated, with Spearman correlation coefficient $\in [0.95,1]$ and Pearson coefficient $\in [0.85,1)$ for $ws>1000$. The corresponding PCA plot for the two first components of the PCA exhibited in Table~\ref{compPCA1} is shown in Figure~\ref{PCA}, where each vertex is colored according to the sector they belong. As expected, peripheral vertices have very low values in the first component and greater dispersion in the second component. 
The plot of clustering coefficient versus degree in Figure~\ref{clust} is very similar to the PCA in Figure~\ref{PCA}.
The PCA plot in Figure~\ref{PCA2}, where all metrics are considered, reflects symmetry metrics relevance for the variance. 
This is shown in Table~\ref{compPCA2}  where the clustering coefficient is only relevant for the third principal component (with contributions from out-degree and out-strength). 
It is concluded that the symmetry-related measurements can be more meaningful in characterizing interaction networks (and their participants) than the clustering coefficient, especially for hubs and intermediary vertices, which are more dispersed in Figure~\ref{PCA2} than in Figure~\ref{PCA}.

@@

    \subsection{Primitive typology: the activity of participants from different sectors}\label{sec:pty}

This work is aimed at finding common characteristics among (email) interaction networks. Analysis involved primary measures observance and a formal criteria for coherent ratios of hub, intermediary and hub sectors. In this process, inspection done by visualizations and raw data manipulations suggests agents typological peculiarities. These are initial observations, which inspired this article and other ongoing research:

@@

This ``primitive typology'', characterized by peripheral, intermediary and hub types, can be further scrutinized using concepts involved in other typologies, such Meyer-Briggs, Pavlov or F-Scale. This has no pretension of being a direct result of numeric analysis, it is a description refinement of the found structure, in typological terms, and coherent with complex networks literature. Although initial, this bridges human and exact sciences in the most pertinent way authors were able to, as is herein considered a result.

Another typology suggested by results is about the networks themselves. In accordance with previous research results~\cite{barabasiEvo}, this and paired articles~\cite{rcText,versinus} reports a dipole in human interaction network types: networks with few and stable agents, and (relative) many threads per amount of messages contrasts with networks with many agents of intermittent activity and (relative) few threads per amount of messages.
The seminal Nature Letter by Palla, Barab{\'a}si and Vicsek~\cite{barabasiEvo} has strong confluence with this work, suggesting that the smaller size of MET community is responsible for the stronger hubs observed.

>>>>

\subsection{Variance prevalence of centrality over symmetry and symmetry over clusterization}\label{prevalence}
We have analyzed the topology of the email networks using standard, well-established metrics of centrality and clustering.
We also introduced symmetry measures because there is evidence that they are relevant in social contexts~\cite{newmanEvolving}.
The contribution from the distinct metrics to network topology is very similar for all the networks considered, and did not vary with time.
This stability in network behavior is remarkable, as can be noticed by the small standard deviations of the contribution of each metric to principal components (from usual Principal Component Analysis~\cite{pca}) along time of Tables~\ref{compPCA0}-\ref{compPCA2}.

@@

Degree and strength are highly correlated, with Spearman correlation coefficient $\in [0.95,1]$ and Pearson coefficient $\in [0.85,1)$ for $ws>1000$. The corresponding PCA plot for the two first components of the PCA exhibited in Table~\ref{compPCA} is shown in Figure~\ref{PCA}, where each vertex is colored according to the sector they belong. As expected, peripheral vertices have very low values in the first component and greater dispersion in the second component. 
The plot of clustering coefficient versus degree in Figure~\ref{clust} is very similar to the PCA in Figure~\ref{PCA}.
The PCA plot in Figure~\ref{PCA2}, where all metrics are considered, reflects symmetry metrics relevance for the variance. 
This is shown in Table~\ref{compPCA2}  where the clustering coefficient is only relevant for the third principal component (with contributions from out-degree and out-strength). 
It is concluded that the symmetry-related measurements can be more meaningful in characterizing interaction networks (and their participants) than the clustering coefficient, especially for hubs and intermediary vertices, which are more dispersed in Figure~\ref{PCA2} than in Figure~\ref{PCA}.

@@

    \subsection{Primitive typology: the activity of participants from Erd\"os sectors}\label{sec:pty}

Visualizations and raw data manipulations suggests agents typological peculiarities. These are initial observations, which inspired this article and other ongoing research~\cite{rcText,versinus}:

@@

This ``primitive typology'', characterized by peripheral, intermediary and hub types, can be further scrutinized using concepts from other typologies, such Meyer-Briggs, Pavlov or F-Scale. This has no pretension of being a direct result of numeric analysis, it is a description refinement of the found structure, in typological terms, and coherent with complex networks literature. Although initial, this bridges human and exact sciences in the most pertinent way authors were able to, as is herein considered a result.

Another typology suggested by results is about the networks themselves. This and paired articles~\cite{rcText,versinus} reports a dipole in human interaction network types: networks with few and stable agents, and (relative) many threads per amount of messages contrasts with networks with many agents of intermittent activity and (relative) few threads per amount of messages.
This network typology is endorsed by the seminal Nature Letter by Palla, Barab{\'a}si and Vicsek~\cite{barabasiEvo}, which suggests that the smaller size of MET community is responsible for the stronger hubs observed.
